---
title: "IPCC AR6 WG1 Ch.6 Fig.19"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Initial steps

Add required libraries and tag scenarios to remove from datasets (per plots.py)

```{r init}
suppressMessages(library(tidyverse))
suppressMessages(library(sf))
suppressMessages(library(tmap))
suppressMessages(library(spData))

# Unused scenarios
unused_scenarios <- c("SSP4-3.4-SPA4", "SSP4-6.0-SPA4", "SSP5-3.4-OS")
```

## Mapping countries to regions

### Preliminary mapping

From the file `country_mapping_ISO-5Regions.csv`, select only the three columns relevant per `plots.py` (adding “Country.Name” to check which countries are mapped where, and for making a map later).

Scroll down for an explanation.

```{r country_mapping}
country_mapping <- 
	read.csv("country_mapping_ISO-5Regions.csv") %>% 
	select(c(Country.Name, current.region.defs, NEW.AR6.Ch6.Fig6.4))

country_mapping

region_mapping <- 
	country_mapping %>%
	group_by(current.region.defs, NEW.AR6.Ch6.Fig6.4) %>%
	summarise(count = n()) %>%
	arrange(current.region.defs)

region_mapping

region_mapping %>% 
	select(!count) %>% 
	summarise(count = n()) %>% 
	arrange(desc(count))
```

**EXPLANATION** Above, you can see the `r nrow(region_mapping)` “current.region.defs” regions of various countries mapped to `r length(unique(region_mapping$NEW.AR6.Ch6.Fig6.4))` different “NEW.AR6.Ch6.Fig6.4” regions.
Evidently, there are some overlaps.
We explore these in the next section.

### Choice of mapping “current.region.defs” to “NEW.AR6.Ch6.Fig6.4”

First we see the breakdown of the number of countries in each of the `r length(unique(region_mapping$NEW.AR6.Ch6.Fig6.4))` “NEW.AR6.Ch6.Fig6.4” regions.

Then we look at which countries are resonsible for there being no many-to-one mapping between the old and new region definitions.

```{r region-splits}
country_mapping %>% 
	select(!current.region.defs) %>% 
	group_by(NEW.AR6.Ch6.Fig6.4) %>% 
	summarise(count = n()) %>% 
	arrange(desc(count))

region_EA <- 
	country_mapping %>% 
	filter(current.region.defs == "Eastern Asia")

region_EA %>% 
	filter(NEW.AR6.Ch6.Fig6.4 == "East Asia")

region_EA %>% 
	filter(!NEW.AR6.Ch6.Fig6.4 == "East Asia")

region_NA <- 
	country_mapping %>% 
	filter(current.region.defs == "North America")

region_NA %>% 
	filter(NEW.AR6.Ch6.Fig6.4 == "Central and South America")

region_NA %>% 
	filter(!NEW.AR6.Ch6.Fig6.4 == "Central and South America")

region_SEA_DP <- 
	country_mapping %>% 
	filter(current.region.defs == "South-East Asia and Developing Pacific")

region_SEA_DP %>% 
	filter(NEW.AR6.Ch6.Fig6.4 == "East Asia")

region_SEA_DP %>% 
	filter(!NEW.AR6.Ch6.Fig6.4 == "East Asia")
```

## Regions on the map

Before showing the regions on a map, we first see if the names of the countries match between the two databases.
Below you can see that they don’t for several countries (and some regions, such as “French Southern and Antarctic Lands”, which are absent in the IPCC database).
For example, “Russian Federation” is “Russia” in the IPCC database, while “eSwatini” is “Swaziland” etc.
Nevertheless, the map below shows the distribution of regions.

```{r world-map}
world_ipcc <- 
	left_join(world, country_mapping, 
			  by = join_by(name_long == Country.Name))

world_ipcc %>% 
	select(name_long, NEW.AR6.Ch6.Fig6.4) %>% 
	filter(is.na(NEW.AR6.Ch6.Fig6.4))

tm_shape(world_ipcc) + 
	tm_fill(col = "NEW.AR6.Ch6.Fig6.4") + 
	tm_borders()
```

## Databases of global and regional emissions

### Global emissions

Below you can see the variables from the database, the number of rows (observations), the models used, the scenarios considered and the breakdown of variables.

```{r emissions-global-data}
emissions_global <- 
	read.csv("ar6-wg1-ch6-emissions-global-data.csv") %>% 
	filter(!Scenario %in% unused_scenarios)

cols_global <- colnames(emissions_global)
cols_global

nrow(emissions_global)

unique(emissions_global$Model)

unique(emissions_global$Scenario)

unique(emissions_global$Region)

unique(emissions_global$Variable)
```

### Regional emissions

The same information as above is shown here, but for regional emissions from the file `ar6-wg1-ch6-emissions-regional-data-5regions.csv`.

```{r emissions-regional-data}
# N.B.: Uses five regions
emissions_regional <- 
	read.csv("ar6-wg1-ch6-emissions-regional-data-5regions.csv") %>% 
	filter(!Scenario %in% unused_scenarios)

cols_regional <- colnames(emissions_regional)
cols_regional

nrow(emissions_regional)

unique(emissions_regional$Model)

unique(emissions_regional$Scenario)

unique(emissions_regional$Region)

unique(emissions_regional$Variable)
```

### Comparing global and regional databases

While the database of global emissions has `r length(cols_global)` variables, that for regional emissions only has `r length(cols_regional)`.
Below we see which variables are missing from the latter.

```{r compare-columns}
janitor::compare_df_cols(emissions_global, emissions_regional) %>% 
	filter(is.na(emissions_regional))
```

### Comparing two CSV files of regional data

The IPCC has provided two seemingly identical databases with different names:
`ar6-wg1-ch6-emissions-regional-data-5regions.csv` and `ar6-wg1-ch6-emissions-regional-data.csv`.
Initially, it seemed like the latter had more granular information about the regions (perhaps mapped to “current.region.defs”), but this is not the case: it has the same region names as the other file.
Further exploration shows some differences between the two files, as shown below.

```{r compare-regional-csv-files}
emissions_current_regions <- 
	read.csv("ar6-wg1-ch6-emissions-regional-data.csv") %>% 
	filter(!Scenario %in% unused_scenarios) 
# %>% 
# 	rename(Current_Region = Region)

unique(emissions_current_regions$Region)

summary(arsenal::comparedf(emissions_current_regions, emissions_regional))
```
## Comparing emission types

Based on the below graph (and the ones in the sections below), it appears that the emissions recorded by just the name of the chemical are the aggregate value of the individual contributing factors.

```{r emission-types}
emissions_regional %>% 
	filter(grepl("CH4", Variable)) %>% 
	filter(Region == "Africa and Middle East") %>% 
	filter(Model == "AIM") %>% 
	filter(grepl("LowNTCF", Scenario)) %>% 
	select(!Model) %>% 
	select(!Scenario) %>% 
	select(!Region) %>% 
	select(!Unit) %>% 
	rename_with(~ str_remove(., "X"), everything()) %>% 
	pivot_longer(
		cols = "1850":"2100",
		names_to = "year",
		values_to = "count"
	) %>% 
	mutate(year = as.Date(year, format = "%Y")) %>% 
	mutate(year = as.numeric(format(year, "%Y"))) %>% 
	filter(year >= 2020) %>% 
	#select(!year) %>% 
	ggplot() + 
	geom_col(aes(count, fct_reorder(Variable, count))) + 
	facet_grid(cols = vars(year))
```


## Reproduction of graphics

First we look at methane emissions in Africa and the Middle East, to compare with the top-left plot from the [original figure](https://www.ipcc.ch/report/ar6/wg1/downloads/figures/IPCC_AR6_WGI_Figure_6_19.png).

```{r methane-ame, warning = FALSE}
emissions_regional %>% 
	filter(Variable == "Emissions|CH4") %>% 
	select(!Variable) %>% 
	filter(Region == "Africa and Middle East") %>% 
	select(!Region) %>% 
	select(!Unit) %>%  
	unite(Model_Scenario, c("Model", "Scenario")) %>% 
	rename_with(~ str_remove(., "X"), everything()) %>% 
	pivot_longer(
		cols = "1850":"2100",
		names_to = "year",
		values_to = "count"
	) %>% 
	mutate(year = as.Date(year, format = "%Y")) %>% 
	mutate(year = as.numeric(format(year, "%Y"))) %>%
	ggplot() +
	geom_point(aes(x = year, y = count, colour = Model_Scenario)) +
	geom_line(aes(x = year, y = count, colour = Model_Scenario)) 
```

Then we look at the RCP-only data, which matches with the grey area in the original figure. It appears that for each year, the minimum and maximum of all four datasets is used to plot the boundaries of the grey area.

```{r methane-ame-rcp-only, warning = FALSE}
emissions_regional %>% 
	filter(Variable == "Emissions|CH4") %>% 
	select(!Variable) %>% 
	filter(Region == "Africa and Middle East") %>% 
	select(!Region) %>% 
	select(!Unit) %>%  
	unite(Model_Scenario, c("Model", "Scenario")) %>% 
	rename_with(~ str_remove(., "X"), everything()) %>% 
	pivot_longer(
		cols = "1850":"2100",
		names_to = "year",
		values_to = "count"
	) %>% 
	filter(grepl("RCP", Model_Scenario)) %>% 
	mutate(year = as.Date(year, format = "%Y")) %>% 
	mutate(year = as.numeric(format(year, "%Y"))) %>%
	ggplot() +
	geom_point(aes(x = year, y = count, colour = Model_Scenario)) +
	geom_line(aes(x = year, y = count, colour = Model_Scenario)) 
```

Below is the faceted breakdown of all emissions by region.
N.B.: Where the original plot shows data for “NMVOC”, it comes from the value of ”Emissions|VOC” in the dataset.
Also note that I cannot get the order of the gases to match the original plot (yet).
The labelling also needs some work, but the plotting itself seems to be OK.

```{r all-emissions-faceted, warning = FALSE}
emissions_type <- c("Emissions|CH4"
				  , "Emissions|BC"
				  , "Emissions|Sulfur"
				  , "Emissions|CO"
				  , "Emissions|NH3"
				  , "Emissions|VOC"
				  , "Emissions|NOx"
				  , "Emissions|OC"
					)

emissions_regional %>% 
	filter(Variable %in% emissions_type) %>% 
	mutate(Variable = as.factor(Variable)) %>% 
	select(!Unit) %>%  
	unite(Model_Scenario, c("Model", "Scenario")) %>% 
	rename_with(~ str_remove(., "X"), everything()) %>% 
	pivot_longer(
		cols = "1850":"2100",
		names_to = "year",
		values_to = "count"
	) %>% 
	mutate(year = as.Date(year, format = "%Y")) %>% 
	mutate(year = as.numeric(format(year, "%Y"))) %>%
	ggplot() +
	geom_point(aes(x = year, y = count, colour = Model_Scenario)) +
	geom_line(aes(x = year, y = count, colour = Model_Scenario)) +
	facet_grid(rows = vars(Variable), cols = vars(Region), scales = "free_y")


# mutate(Variable = fct_relevel(Variable),
# 		   levels = c("Emissions|CH4"
# 		   		    , "Emissions|BC"
# 		   		    , "Emissions|OC"
# 		   		    , "Emissions|Sulfur"
# 		   		    , "Emissions|NOx"
# 		   		    , "Emissions|CO"
# 		   		    , "Emissions|VOC"
# 		   		    , "Emissions|NH3"
# 		   		    ))
	
```


